---
title: 'Final Project: Data Cleaning'
author: "Michael Pascutoi"
date: "12/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Read In and Summarize Data
```{r Libraries}
library(tidyverse)
library(dplyr)
library(stringr)
library(fastDummies)

```

```{r Read in the Data}
imdb_data<-read.csv("./movie_ratings.csv",header=TRUE, na.strings=c("","NA"))

```


```{r Initial Summary and Cleaning}
#General Data basics
colnames(imdb_data)
dim(imdb_data)
head(imdb_data)
str(imdb_data)


#Remove unncessary characters from numeric data and make 'year' a numeric
imdb_data$budget = as.numeric(gsub("\\$", "", imdb_data$budget))
imdb_data$usa_gross_income=as.numeric(gsub("\\$", "", imdb_data$usa_gross_income))
imdb_data$worlwide_gross_income=as.numeric(gsub("\\$", "", imdb_data$worlwide_gross_income))
imdb_data$year <- as.numeric(imdb_data$year) 


#Production Company had multiple names, which we had to condense
imdb_data$production_company <- recode(imdb_data$production_company,
                              "20th Centruy Fox Film Corporation"="Twentieth Century Fox",
                              "Fox 2000 Pictures" = "Twentieth Century Fox")


```

##Creating the Data Set
```{r}
#Removed films not made in the US due to issues with currency in dependent variable
imdb_data <- imdb_data  %>%
  filter(country=="USA")
```


```{r Preparing The Data for Dummy Separation}
#Remove films that did not have budget or income information
imdb_data<-imdb_data[!is.na(imdb_data$budget) & !is.na(imdb_data$usa_gross_income),] 

##Separate cast and crew information into separate columns, and add placeholder columns for production companies
imdb_data <- imdb_data %>%
  separate(col=actors,into=c("actor1","actor2","actor3","actor4","actor5"), sep="," ) %>%
  separate(col=director, into=c("director1","director2"), sep = ",") %>%
  separate(col=writer, into=c("writer1","writer2","writer3"), sep=",") %>%
  separate(col=genre, into = c("genre1","genre2","genre3"), sep=",")  %>%
  add_column(Columbia_production=NA,
             Universal_production = NA,
             Fox_production = NA,
             Paramount_production=NA,
             Warner_production = NA,
             Disney_production = NA,
             Lionsgate_production = NA,
             MGM_production = NA,
             Weinstein_production =NA,
             Pixar_production = NA)
```

#Production Company Information
```{r}
#Create the dummy variables for the production companies
imdb_data$Columbia_production<-ifelse(grepl("Columbia", imdb_data$production_company), 1, 0)  
imdb_data$Universal_production<-ifelse(grepl("Universal", imdb_data$production_company), 1, 0)  
imdb_data$Fox_production<-ifelse(grepl("Century Fox", imdb_data$production_company), 1, 0)  
imdb_data$Paramount_production<-ifelse(grepl("Paramount", imdb_data$production_company), 1, 0)  
imdb_data$Warner_production<-ifelse(grepl("Warner Bros", imdb_data$production_company), 1, 0)  
imdb_data$Disney_production<-ifelse(grepl("Walt Disney P", imdb_data$production_company), 1, 0)  
imdb_data$Lionsgate_production<-ifelse(grepl("Lionsgate", imdb_data$production_company), 1, 0)  
imdb_data$MGM_production<-ifelse(grepl("MGM", imdb_data$production_company), 1, 0)  
imdb_data$Weinstein_production<-ifelse(grepl("Weinstein", imdb_data$production_company), 1, 0)  
imdb_data$Pixar_production<-ifelse(grepl("Pixar", imdb_data$production_company), 1, 0)  
 
```


```{r}
#Create dependent variable through existing information in data set
imdb_dataset<- imdb_data  %>% 
  select(-c("X","imdb_title_id","title","date_published","description")) %>%
  mutate(domestic_profit = usa_gross_income - (budget*1.5), 
       worldwide_profit = worlwide_gross_income - (budget*1.5)) %>%
  mutate(domestic_profitability = domestic_profit/budget,
         worldwide_profitability = worldwide_profit/budget, 
         profit = case_when(worldwide_profit > 0 ~ 1,
                            worldwide_profit <= 0  ~0))

#Add variable for number of langauges in film and dummy variable for presence of language besides English 
imdb_dataset <- imdb_dataset %>%
  mutate(language_number = count.fields(textConnection(language), sep = ",")) %>%
  mutate(language_bool = case_when(language_number > 1 ~ 1,
                                   language_number == 1 ~ 0))

##Remove NA's from dataset and change variables to factors or numerics for testing
imdb_dataset[is.na(imdb_dataset)] <- ""  
imdb_dataset$profit <- as.factor (imdb_dataset$profit)
imdb_dataset$metascore <-as.integer(imdb_dataset$metascore)

```
##Load Dummy Variables created from director and actor information
#Note: These dummy variables were produced with the help of code provided by Professor Martin Barron

#Create Dummy Variables for film director
```{r}
dc1 <- dummy_cols(imdb_dataset$director1, remove_selected_columns = TRUE)
dc2 <- dummy_cols(imdb_dataset$director2, remove_selected_columns = TRUE)

dc2[is.na(dc2)] <- 0

temp <- cbind.data.frame(dc1, dc2)

directors <- unique(c(imdb_dataset$director1, imdb_dataset$director2))
res_db <- as.data.frame(matrix(NA, nrow = nrow(imdb_dataset), ncol = length(directors)))
for(i in 1:length(directors)){
  col <- paste(".data_", directors[i], sep = "")
  temp <- cbind.data.frame(dc1[,which(names(dc1) == col)],
                           dc2[,which(names(dc2) == col)])
  res_db[,i] <- rowSums(temp)
}
# Fix names
names(res_db) <- directors
colnames(res_db) <-paste(colnames(res_db),"director", sep="_")
names(res_db) <- gsub(" ", "_", names(res_db))
names(res_db) <- gsub("^_", "", names(res_db))

director_db <- res_db
```

#Create Dummy Variables for Actors/Actresses
```{r}
da1 <- dummy_cols(imdb_dataset$actor1, remove_selected_columns = TRUE)
da2 <- dummy_cols(imdb_dataset$actor2, remove_selected_columns = TRUE)
da3 <- dummy_cols(imdb_dataset$actor2, remove_selected_columns = TRUE)
da4 <- dummy_cols(imdb_dataset$actor4, remove_selected_columns = TRUE)
da5 <- dummy_cols(imdb_dataset$actor5, remove_selected_columns = TRUE)


da1[is.na(da1)] <- 0
da2[is.na(da2)] <- 0
da3[is.na(da3)] <- 0
da4[is.na(da4)] <- 0
da5[is.na(da5)] <- 0

actors <- unique(c(imdb_dataset$actor1, imdb_dataset$actor2, imdb_dataset$actor2,
                   imdb_dataset$actor4, imdb_dataset$actor5))

res_db_1 <- as.data.frame(matrix(NA, nrow = nrow(imdb_dataset), ncol = length(actors)))
for(i in 1:length(actors)){
  act <- paste(".data_", actors[i], sep = "")
  temp <- cbind.data.frame(da1[, which(names(da1) == act)],
                           da2[, which(names(da2) == act)],
                           da3[, which(names(da3) == act)],
                           da4[, which(names(da4) == act)],
                           da5[, which(names(da5) == act)])
  res_db_1[,i] <- rowSums(temp)
}
names(res_db_1) <- actors

colnames(res_db_1) <-paste(colnames(res_db_1),"actor", sep="_")
names(res_db_1) <- gsub(" ", "_", names(res_db_1))
names(res_db_1) <- gsub("^_", "", names(res_db_1))

actor_db <- res_db_1
save(actor_db, file ="actor_res.rda")

```


#Create Dummy Variables for film genre
```{r}
#Remove excess spaces within the variables
imdb_dataset$genre1 <- recode(imdb_dataset$genre1,
                     " Music"="Music",
                     " Romance"= "Romance",
                     " Drama"= "Drama",
                     " Family" = "Family",
                     " History" = "History",
                     " Comedy" = "Comedy",
                     " Mystery" = "Mystery",
                     " Fantasy" = "Fantasy",
                     " Crime" = "Crime",
                     " Film-Noir" = "Film-Noir",
                     " Musical"= "Musical",
                     " Thriller" = "Thriller",
                     " Adventure" = "Adventure",
                     " War" = "War",
                     " Western" = "Western",
                     " Biography" = "Biography",
                     " Sci-Fi" = "Sci-Fi",
                     " Music" = "Music",
                     " Horror" = "Horror",
                     " Sport" = "Sport",
                     " Action" = "Action")

imdb_dataset$genre2 <- recode(imdb_dataset$genre2,
                              " Music"="Music",
                              " Romance"= "Romance",
                              " Drama"= "Drama",
                              " Family" = "Family",
                              " History" = "History",
                              " Comedy" = "Comedy",
                              " Mystery" = "Mystery",
                              " Fantasy" = "Fantasy",
                              " Crime" = "Crime",
                              " Film-Noir" = "Film-Noir",
                              " Musical"= "Musical",
                              " Thriller" = "Thriller",
                              " Adventure" = "Adventure",
                              " War" = "War",
                              " Western" = "Western",
                              " Biography" = "Biography",
                              " Sci-Fi" = "Sci-Fi",
                              " Music" = "Music",
                              " Horror" = "Horror",
                              " Sport" = "Sport",
                              " Action" = "Action")

imdb_dataset$genre3 <- recode(imdb_dataset$genre3,
                              " Music"="Music",
                              " Romance"= "Romance",
                              " Drama"= "Drama",
                              " Family" = "Family",
                              " History" = "History",
                              " Comedy" = "Comedy",
                              " Mystery" = "Mystery",
                              " Fantasy" = "Fantasy",
                              " Crime" = "Crime",
                              " Film-Noir" = "Film-Noir",
                              " Musical"= "Musical",
                              " Thriller" = "Thriller",
                              " Adventure" = "Adventure",
                              " War" = "War",
                              " Western" = "Western",
                              " Biography" = "Biography",
                              " Sci-Fi" = "Sci-Fi",
                              " Music" = "Music",
                              " Horror" = "Horror",
                              " Sport" = "Sport",
                              " Action" = "Action")

#create dummy columns for genres, making sure that films with multiple genres can have all listed
gdum1<-dummy_cols(imdb_dataset$genre1, remove_selected_columns = TRUE)
gdum2<-dummy_cols(imdb_dataset$genre2, remove_selected_columns = TRUE)
gdum3<-dummy_cols(imdb_dataset$genre3, remove_selected_columns = TRUE)

#Convert NAs to 0
gdum2[is.na(gdum2)] <- 0
gdum3[is.na(gdum3)] <- 0

#Combine dummy variables into one data frame
gdum_frame <- cbind.data.frame(gdum1, gdum2, gdum3)

#Find all unique genre names
genres <- unique(c(imdb_dataset$genre1, imdb_dataset$genre2, imdb_dataset$genre3))

#Create database for all genres
genre_db <- as.data.frame(matrix(NA, nrow=nrow(imdb_dataset), ncol=length(genres)))

#Combine dummy variable frame with genre data frame to create a set data base
for (i in 1:length(genres)) {
  col <- paste(".data_", genres[i], sep = "")
  gdum_frame <- cbind.data.frame(gdum1[,which(names(gdum1) == col)],
                           gdum2[,which(names(gdum2) == col)],
                           gdum3[,which(names(gdum3) == col)])
  genre_db[,i] <- rowSums(gdum_frame)
  
}
#Apply genre names to data frame
names(genre_db) <- genres

#Summary of information for genre_db
summary(colSums(genre_db))
plot(density(colSums(genre_db)))



```
##Combine all dbs into one data frame

```{r}
total_data_imdb <- cbind(imdb_dataset, director_db, actor_db, genre_db)

save(total_data_imdb, file="./imdb_data_cleaned.rda")

```


##Reading in data for final formatting
```{r}
rm(list = ls())
load("./imdb_data_cleaned.rda")
```

##Creating list of movies based on Video games from Wikipedia
```{r}
library(httr)
library(XML)

url <- "https://en.wikipedia.org/wiki/List_of_films_based_on_video_games"

r <- GET(url)

doc <- readHTMLTable(doc=content(r, "text"))
vg_movies <- c(doc[[1]][2:43,'V1'], doc[[2]][2:5,'V1'])
```

##Adding a new binary column for video game movies
```{r cleaning and adding is_vg}
#names(total_data_imdb) <- gsub(" ", "_", names(total_data_imdb))
#names(total_data_imdb) <- gsub("^_", "", names(total_data_imdb))
for(title in total_data_imdb){
  total_data_imdb$is_vg <- ifelse(total_data_imdb$original_title %in% vg_movies, 1, 0)
}
```


##Removing duplicate columns and varibales that are not independent from the predictor.
```{r filetering out columns}
dup_names <- !duplicated(names(total_data_imdb))
summary(dup_names)
total_data_imdb2 <- total_data_imdb[,dup_names]
summary(duplicated(names(total_data_imdb2)))

dtypes <- sapply(total_data_imdb2, class)
numeric_index <- dtypes %in% c('numeric', 'integer', 'factor')

total_data_num <- total_data_imdb2[,numeric_index]


rm_cols <- c("original_title",
              "year",
              "genre1",
              "genre2",
              "genre3",
              "country",
              "language",
              "director1",
              "director2",
              "writer1",
              "writer2",
              "writer3",
              "production_company",
              "director", 
              "actor",
              "actor1",
              "actor2",
              "actor3",
              "actor4",
              "actor5",
              "avg_vote",
              "votes",
              "usa_gross_income",
              "worlwide_gross_income",
              "metascore",
              "reviews_from_users",
              "reviews_from_critics",
              "weighted_average_vote",
              "total_votes",
              "mean_vote",
              "median_vote",
              "votes_10",
              "votes_9",
              "votes_8",
              "votes_7",
              "votes_6",
              "votes_5",
              "votes_4",
              "votes_3",
              "votes_2",
              "votes_1",
              "allgenders_0age_avg_vote",
              "allgenders_0age_votes",
              "allgenders_18age_avg_vote",
              "allgenders_18age_votes",
              "allgenders_30age_avg_vote",
              "allgenders_30age_votes",
              "allgenders_45age_avg_vote",
              "allgenders_45age_votes",
              "males_allages_avg_vote",
              "males_allages_votes",
              "males_0age_avg_vote",
              "males_0age_votes",
              "males_18age_avg_vote",
              "males_18age_votes",
              "males_30age_avg_vote",
              "males_30age_votes",
              "males_45age_avg_vote",
              "males_45age_votes",
              "females_allages_avg_vote",
              "females_allages_votes",
              "females_0age_avg_vote",
              "females_0age_votes",
              "females_18age_avg_vote",
              "females_18age_votes",
              "females_30age_avg_vote",
              "females_30age_votes",
              "females_45age_avg_vote",
              "females_45age_votes",
              "top1000_voters_rating",
              "top1000_voters_votes",
              "us_voters_rating",
              "us_voters_votes",
              "non_us_voters_rating",
              "non_us_voters_votes",
              "domestic_profit",
              "worldwide_profit",
              "domestic_profitability",
              "worldwide_profitability",
              "Var.11971"  
             )


#Removing problematic columns
total_data_num <- total_data_num[, which(!names(total_data_num) %in% rm_cols)]

#dim(total_data_imdb)
#dim(total_data_num)
```

##Filtering out the varibales that have low number of occurences.
```{r}
col_weights <- colSums(Filter(is.numeric, total_data_num))

narrow_cols <- c(names(col_weights[col_weights >= 7]),
                 "is_vg",
                 
                 "Ruben_Fleischer_director",
                 "Tom_Holland_actor",
                 #"Mark_Wahlberg_actor",
                 #"Antonio_Banderas_actor",
                 
                 "profit"
                 )

total_data <- total_data_num[,narrow_cols]
#dim(total_data)
```

##Saving final formatted file for model training
```{r}
save(total_data, file = 'total_data.rda')
```
